Embodied Neuromorphic Benchmarks

### 2015-08-10 paper/paper.tex
adaptive control

\section{Embodied Benchmark 1: Adaptive Motor Control}

>>>>

\section{A Benchmark: Adaptive Motor Control}

To demonstrate this approach to creating embodied neuromorphic benchmarks,
we now consider a basic control task.  Suppose we have a system with a number
of joints $q$ and we want to send an output $u$ to those motors such that
the joints move to a particular desired position $q_d$.  Our only output is
the signal $u$ (one for each motor) and our only input is the current position
of each motor $q$.

The simplest controller for such a situation is a P (proportional) controller,
where $u=K_p(q_d - q)$.  This is often supplemented with a derivative term,
which helps to slow the system down as it approaches the desired position,
thus avoiding overshooting and oscillation ($u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$),
leading to the standard PD controller.  Both $K_p$ and $K_d$ are constants that can be
hand-tuned to particular situations.

However, this controller has difficulty in the presence of significant external
forces.  For example, consider a single motor controlling the angle of a single
arm.  If the arm is held out to the side, gravity acting on the mass of the
arm itself will pull the arm downward.  Thus to hold the arm still at a
particular $q_d$ will require the controller to apply a force to counteract
gravity.  Since the PD controller always produces an output $u=0$ when $q=q_d$,
it cannot compensate for this. [TODO: add diagram]

The standard solution to this problem is to add an integral term ($K_i \int{q_d-q dt}$)
to the controller, making it a PID controller.  The idea here is that as the difference between where it is and
where we want it to be accumulates over time, the $K-i$ term gradually increases
how much extra force is being applied until it is large enough to counteract
the external force of gravity (or whatever other external forces are present).
However, this approach has great difficulty when $q_d$ changes, since the
external force due to gravity changes depending on the position of the arm $q$.
The controller ends up having to use the accumulated integral to ``relearn'' the
correct amount of extra force needed every time $q_d$ changes.

In some robotics applications, the solution to this problem is to mathematically
analyze the geometry and mass of the system and compute exactly how much extra
force is needed.  In this particular case, the answer is straight-forward,
in that the extra torque due to gravity is $\tau=m l g sin(q) / 2$, where
$m$ is the mass of the arm, $l$ is the length, and $g$ is $9.8m/s^2$.  If the
force applied by the motor is linear in $u$, then we could simply compute this
value and add it to our controller's output.  However, this assumes a perfectly
even distribution of weight in the arm, ignores momentum and other forces, and 
gets much more complex as more joints are added.  Furthermore, if this
initial computation is slightly off, or if details of the system change,
there is no way to adjust this compensation.

Fortunately, there is an adaptive solution to this problem, and it is one that
fits well with neuromorphic hardware.  In \cite{Slotine1987}, Slotine shows that if you
express the influence of these other external forces as $\tau=Y(q) \omega$
(where $Y(q)$ is a fixed set of functions of $q$, such as $sin(q)$, and \omega
is a vector of weights, one for each function in $Y$), then you can learn
to compensate for these external forces by using the learning rule $\Delta \omega = \alpha Y(q) u$,
where $u$ is the basic PD control signal.

Importantly, as pointed out in [ref] and [ref], rather than making explicit
assumptions about the exact functions that should be in $Y(q)$, we can use
a neural network approach where each neuron is a different function of $q$.  
As long as there is enough hetereogenetity (i.e. as long as the neural activity
forms a basis space that spans the desired space of functions), then the
learning rule will continue to work.

This then suggests an explicit neuromorphic benchmark.  The input to the
neuromorphic hardware is $q$, the system state.  This input is fed to each
neuron such that each neuron produces some output behaviour that is based on
this input.  Since $q$ will be multi-dimensional (if there is more than one
joint), we may give each neuron a random weighting of each $q$ value ($J_i=e_i \cdot q$, 
where $J_i$ is the input to neuron $i$, and $e_i$ is a randomly chosen vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities}).  Given this input, the neurons will produce some output $A$.
we now form a weighted sum of these outputs $Ad$, where $d$ is a matrix (number of neurons by number of elements in $q$)
that is initially all zeros.

To use this controller, we add its output to that of the standard PD controller.
That is, the standard controller has $u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$,
and so our actual output to the motor is $u + Ad$.  We then apply a learning
rule on $d$ such that $\Delta d = \alpha A \times u$.

Notice that we can think of this system as a three-layer neural
network, where the input and output layers are linear.  The first layer is $q$, the input state, one value for each joint.  The ``hidden'' layer
is $A$, the activity of a large number of neurons.  The output layer again has
one value per joint, and is the extra added signal to apply to the motors, $Ad$.
Given that this is such a canonical example of the use of neural networks, we
hope that the majority of neuromorphic hardware is flexible enough to implement
exactly this model.

\subsection{Online versus offline learning}

The one major step here that does not exist in a lot of neuromorphic hardware is
the ability to update the weights $d$.  For hardware that does have a built-in
learning rule, this rule is at least of a very common form, where the weight
update from a neuron is proportional to the activity of that neuron and an
external error signal.  This makes it an instance of the ubiquitous delta rule.

However, if the neuromorphic hardware being benchmarked does not have the
ability to update weights online using a learning rule of this form, then there
are two solutions.  First, the multiplication by $d$ could be done on the
output from the neuromorphic hardware.  There has to be some system to take
the neural output from the hardware and send it to the motor (or the simulation
of the motor).  Instead of outputing the result of $Ad$, the hardware could
output $A$ (the activity of all the neurons), and the interface to the motor
could be responsible for doing the multiplication by $d$ and updating $d$
according to the learning rule.

Alternatively, we can use offline learning.  That is, rather than updating
the weights $d$ all the time, we simply record $A$ and $u$, and then after
a period of time stop the controller, compute the sum total of the changes
to $d$, load the new value of $d$ onto the neuromorphic hardware, and then
start the controller again.
